Search for Spatially Extended \fermi-LAT Sources Using Two Years of Data

### 2012-04-25 extended_catalog/external_review_reply/replies_to_review_comments.tex

\documentclass{article}

\usepackage{url}
\usepackage{color}

\newenvironment{referee}
{\begin{quote}\color{red}}
  {\end{quote}}

\newenvironment{reply}
  {}
  {}

\title{Reply to Referee Report}

\begin{document}

\maketitle

\begin{reply}
Thank you for your careful review of the text. We discuss below the
changes we have made incorporating your feedback. To make the changes more visible,
in the new draft we have bolded and colored red the parts that are new or have changed.
\end{reply}

\begin{referee}
Major issues :

The authors use a uniform disk model for fitting the extended sources.
However many sources show a complex morphology (for example, not radial symmetric)
They should quantify how the results they are presenting depend on the actual shape of the sources.
\end{referee}

\begin{reply}
In the new draft, Section 3 (Validation of Analysis Method) has been
divided into two sections. The original section is now section 3.1
(now called Statistical Significance Testing).

There is a new section 3.2 called ``Assessing the Bias Due to Incorrect
Spatial Models''. In this section, we address your question by performing
a dedicated Monte Carlo study of W44 simulated with a ring-type spatial
model. By fitting the source with different spatial models, we asses
this bias and find it to be very small.

This new section requires two new figures (Figure 6 and Figure 7). Also,
in section 2.4 (Comparing Source Size) we add a line of text referring
to this new section. Finally, there is a new sentance in the introduction
of the paper which references the new text.
\end{reply}

\begin{referee}
Most of the new extended sources found in this work lie on the galactic
plane at very low latitudes. Why the simulations used to validate the
analysis (par. 3) do not include the diffuse Galactic background? Can
the structures of this diffuse background affect the analysis results?
\end{referee}

\begin{reply}
In section 3.1 (which was previously just section 3 and is called
Statistical Significance Testing), we have added new text describing a
follow up Monte Carlo study performed in the presence of the galactic
plane. In addition, this new text required updating Table 1 (Monte Carlo
Spectral Parameters) to include the new simulation parameters. There is also
a new figure (Figure 5) which shows the distribution of TSext for these
new simulations.
\end{reply}

\begin{referee}
An important result of this work is that some 2FGL sources could be fake due to the residuals induced by modeling the extended sources as point-like. It may be worth having a table that lists the sources that are not longer needed in the modeling of the sky (alternatively this information could be included in table 4). They must also be mentioned in the captions of figures 16, 17, 19, 21 and 22.
\end{referee}

\begin{reply}
We have added a new table (Table 6 in the current draft) which contains
this information. There are paragraphs in section 9 (Extended Source
Search Method) and Section 10 (New Extended Sources) which refer to this
table. We have also updated the specified captions to list names of 2FGL
sources that were removed.
\end{reply}

\begin{referee}
Minor issues :

Par. 3, eq 8 : The authors should explain the symbols used
\end{referee}

\begin{reply}
The symbols are now described after the equation.
\end{reply}

\begin{referee}
Par. 7, row 386 : W30 is cited 2 times
\end{referee}

\begin{reply}
Fixed
\end{reply}

\begin{referee}
  Par. 10.3, rows 650-651 : The authors should explain the relation between $r_{68}$ and $\sigma$
\end{referee}

\begin{reply}
  The relation between $r_{68}$ and $\sigma$ was already included in the text
in section 2.4 (Comparing Source Sizes). But we restructured the
text in that section to bring to prominence the relevant equations.
\end{reply}

\begin{referee}
Par. 10.4, row 674 : The authors should quantify the statistical
improvement using one single extended source instead of 2 pointlike
\end{referee}

\begin{reply}
The numbers are now included in the text
\end{reply}

\begin{referee}
Par. 10.6, row 720 : The authors should explain better how you distinguish
if a source is ``physically separate'' or not
\end{referee}

\begin{reply}
The text has been clarified
\end{reply}

\begin{referee}
Par. 10.8, row 768 : The authors should quantify the statistical
improvement using one single extended source instead of 2 pointlike
\end{referee}

\begin{reply}
The numbers are included in the text
\end{reply}

\begin{referee}
Fig. 2 : Is the y axis logarithmic ? Please clarify.
\end{referee}

\begin{reply}
The legend now says ``log(PDF) [Arbitrary Units]''
\end{reply}

\begin{referee}
Fig. 14 : The authors should specify the spectral index of the two plotted power-laws (or else remove them).
\end{referee}

\begin{reply}
The spectral indices have been added to the caption.
\end{reply}

\begin{referee}
++++++++++++++++
Report on statistical methodology

Search for Spatially Extended Fermi-LAT Sources Using Two Years of Data

J. Lande et al. (ApJ86244)

a) line 98 states: In the pointlike package, ``The data are binned
spatially, using a HEALPix pixelization, and spectrally (G ́orski et
al. 2005) and the likelihood is maximized over all bins in a region.'' Why
is any binning used? Maximum likelihood estimation (MLE) is best used
on exact photon positions, and binning is not required (as it is with
older chi-square-minimiization techniques). Binning is ill-advised as
it results in a loss of information, and sometimes in a bias. As this
assumption is built into pointlike, perhaps it cannot be changed by the
authors for their analysis.
\end{referee}

\begin{reply}
The Fermi LAT analysis package gtlike has an unbinned likelihood mode
which is indeed slightly more precise. But running an unbinned analysis
is very slow because Fermi collects a large number of photons, many of
which do not contribute very much to the detection of a source. gtlike
in binned mode is much faster and is therefore the most common tool
for LAT analysis. Because the LAT collaboration has extensively tested
binned gtlike for spectral analysis, we do not expect it to introduce
any large bias in our spectral analysis result.

But even binned gtlike is prohibitively slow for the extended source
spatial fitting described in this paper. That is why we developed this
new functionality inside of pointlike. The validations of pointlike
presented in this paper allow us to be confident in our spatial analysis results.
\end{reply}

\begin{referee}
b) The likelihood ratio test (LRT, eqns 1 \& 9) is a classic technique of
MLE that has been largely replaced by statisticians (since the 1980s) by
closely related measures that take into account the complexity of models
and size of the dataset. The LRT has serious deficiencies: it tends
to accept the more complex model as sample size (i.e., source counts)
increases (or equivalently, it assesses only bias and not variability),
and it is inapplicable when parameters (e.g. source sizes) are near zero
(discussed in the ApJ by Protassov et al. 2002). I also am not confident
that the chi-squared approximation in eqn 8 is correct (please refer to
a statistics book not an old ApJ article, or consult a statistician).

Model selection today is instead guided by `penalized' measures,
and the most widely used is the Bayesian Information Criterion
(BIC,\url{http://en.wikipedia.org/wiki/Bayesian_information_criterion}). It
is very easy to compute, essentially an offset from the LRT scaled to
the log of the counts involved. The issues are discussed in detail in
the volume ``Model Selection and Multi-Model Inference'' (K.P. Burnham
\& D. Anderson, 2nd ed, 2002); see also lectures (especially \#14) at
\url{http://myweb.uiowa.edu/cavaaugh/ms_seminar.html}.

I guess you could argue that your Monte Carlo experiments (Figs 3-4) take
the flux and degrees of freedom into account for the LRT statistic. But
if you want your method to be widely used, it should not rest on a weak
and obsolete statistical foundation. Altogether, I recommend moving to
the BIC. Your approach of Monte Carlo testing of the significance of
LRT values should also work for BIC values.
\end{referee}

\begin{reply}
Concerning the extension significance test, we extensively validated it
against Monte Carlo simulations. Furthermore, the high energy astrophysics
community has a lot of experience with the test statistic (TS) applied to
nested models. So we are very confident that we can apply it to the actual
data.

Concerning the test comparing two point-like sources to one extended
source, we find the desire for a Bayesian test more compelling since the
models are not nested. We considered the BIC test but found that it would
not work well for Fermi LAT. We alternately considered the AIC test and
found that it is equivalent to the test we actually performed. We have
added a new paragraph to Section 5 (Testing Against Source Confusion)
that describes the Bayesian interoperation of the test we performed.
\end{reply}

\begin{referee}
Finally, I can suggest that the lead author discuss the issue with experts
at his institution, as Stanford's Dept. of Statistics is universally
recognized as the best in the nation. They might provide a balanced
judgment whether the BIC is substantially better than the LRT in the
situation under study.
\end{referee}

\begin{reply}
Thanks for the suggestion.  In addition to the above changes, we have
fixed the bibliographical information on the LAT publication of Vela Jr.
\end{reply}

\end{document}

### 2012-04-25 extended_catalog/paper/extended_search.tex
updated draft

 For these reasons, in Section~\ref{analysis_methods_section}
we present a new systematic method for analyzing spatially extended
LAT sources.  
\newtext{
In Section~\ref{monte_carlo_validation}, we demonstrate
that this method can be used to test the statistical significance of the
extension of a LAT source and in Section~\ref{bias_wrong_spatial_model}
we asses the bias that is introduced by using an incorrect spatial model.
In Section~\ref{extension_sensitivity}
we calculate the LAT detection threshold to resolve the extension
of a source.}
In Section~\ref{dual_localization_method}, we
study the ability of the LAT to 
distinguish between a single extended source and unresolved closely-spaced point-like sources
In Section~\ref{test_2lac_sources}, we further demonstrate that our
detection method does not misidentify point-like sources as being
extended by testing the extension of active Galactic nuclei (AGN)
believed to be unresolvable.  In Section~\ref{validate_known},
we systematically reanalyze the twelve extended sources included
in the 2FGL catalog and in Section~\ref{systematic_errors_on_extension}
we describe a way to estimate systematic errors on the measured extension of a source.
In Section~\ref{extended_source_search_method}, we
describe a search for new spatially extended LAT sources. Finally,
in Section~\ref{new_ext_srcs_section} we present the detection of the
extension of nine spatially extended sources that were reported in the 2FGL catalog
but treated as point-like in the analysis.  Two of these sources have been previously analyzed in
dedicated publications.

@@

For sources that are comparable in size to the PSF,
the differences in the PDF for
different spatial models are lost in the noise and the LAT is not sensitive
to the detailed spatial structure of these sources.  
\newtext{In section \ref{bias_wrong_spatial_model}, we perform a dedicated Monte Carlo simulation
to asses possible bias due to incorrectly modeling the spatial structure
of an extended source.
In our search for extended sources we use only a radially-symmetric uniform
disk spatial model. Unless otherwise noted,
we quote the radius to the edge ($\sigma$) as the size of the source.}

@@

We evaluated the false detection probability for the measurement of source extent
by fitting the extensions of point-like sources.  \cite{mattox_egret}
discuss that the TS distribution for a likelihood ratio test
on the existence of a source at a given position is
\begin{equation}\label{ts_ext_distribution}
  P(\ts)=\onehalf(\chi^2_1(\ts)+\delta(\ts)),
\end{equation}
\newtext{
where $P(\ts)$ is the probability of getting a particular value of TS,
$\chi^2_1$ is the chi-squared distribution with one degree of freedom, and
$\delta$ is the Dirac delta function.
}
The particular form of Equation \ref{ts_ext_distribution} is due to the
null hypothesis (source flux $\Phi=0$) residing on the edge of parameter
space and the model hypothesis adding a single degree of freedom.
It leads to the often quoted result $\sqrt{TS}=\sigma$, where 
$\sigma$ here refers to the significance of the detection. It is plausible
to expect a similar distribution for the TS in the test for
source extension since the same conditions apply (with the source flux
$\Phi$ replaced by the source radius $r$ and $r<0$ being unphysical).
To verify Equation~\ref{ts_ext_distribution}, we evaluated the
empirical distribution function of \tsext computed from simulated sources.

@@

\newtext{
These simulations spanned a large range of fluxes 
for two different energy ranges assuming an isotropic background.
We performed a second simulation to assess any effects that could
be introduces by performing this test in the presence of a highly
structured background.

Our simulation setup was the same as above except that the sources were
simulated on top of and analyzed assuming the presence of the standard
Galactic diffuse and isotropic background models used in 2FGL.  In our
simulations, we selected our sources to have a random position on the sky
such that the absolute value of the galactic latitude of the source was
less than 5\degree. This probes the most structured areas of the galaxy.

To limit the number of tests, we selected only one flux
level for each of the four spectral indices and we perform
this test only in the 1 \gev to 100 \gev energy range. 
The fluxes were selected so that $\ts\sim50$ which is around 
the significance level for which on might
expect to be able to spatially resolve the source. 
We only tested these relatively dim sources because they are most likely
to be affected by the background.

Because the Galactic diffuse emission is highly structure, the point-source
sensitivity can vary significantly across the galactic plane.
To account for this, we scaled the flux (for a given spectral index)
so that the source always has approximately the same signal-to-noise ratio:
\begin{equation}
  \label{scale_flux_by_background}
  F(\vec{x}) = F(\text{GC}) \times \left(
  \frac{B(\vec{x})}{B(\text{GC})}\right)^{1/2}.
\end{equation}
Here, $\vec{x}$ is the position of the simulated source, $F$ is the integral
flux of the source from 100 \mev to 100 \gev, $F(\text{GC})$
is the same quantity if the source were at the galactic center, $B$
is the integral of the Galactic diffuse and isotropic emission
from 1 \gev to 100 \gev and $B(\text{GC})$ is the same quantity
if the source were at the galactic center.  Table~\ref{ts_ext_num_sims}
contains $F(\text{GC})$ and the average value of \ts for the four spectral models.
For each spectrum, we performed $\sim90,000$ simulations.

Figure~\ref{tsext_plane_plot} shows the cumulative density
of \tsext compared to the $\chi^2_1/2$ distribution of
Equation~\ref{ts_ext_distribution}. For small values of \tsext,
there is good agreement between the simulations and
theory.  For the highest values of \tsext, there is possibly
a small discrepancy, but limited statistics makes this hard to quantify.
}

@@

\newtext{
We performed a Monte Carlo study to asses the bias that is introduced by
incorrectly modeling the spatial structure of an extended source. This
is important because any assumed spatial model is just an approximation
to the true shape of a source.

To assess this systematic, we simulated the spatially extended ring-type SNR W44.
We selected W44 because it is the most significant extended source detected by the LAT 
that has a non-radially symmetric photon-distribution \citep{w44}.

W44 was simulated with a power-law spectral model with an integral flux
of $7.12\times10^{-8}$ \phflux in the energy range from 1 \gev to 100
\gev and a spectral index of 2.66.
The flux and spectral index of this source
were taken from our analysis of W44 which is described in Section~\ref{validate_known}.

W44 was simulated with the elliptical ring spatial model described in
\cite{w44}.  Because that publication does not provide the numbers describing
the parameters of the ellipse, we summarize them here for completeness. The 
ellipse has a semi-major axis of 0\fdg3 and a semi-minor axis of 0\fdg19. The
ellipse predicts emission only for the outer 0.75\% of the shape and
the has a position angle of $147\degree$ measured east of celestial north.

We used a simulation setup similar to that described in
Section~\ref{monte_carlo_validation}, but the source was simulated on
top of and analyzed assuming the presence of the standard Galactic diffuse and
isotropic background model used in 2FGL.  This simulation was
in the 1 \gev to 100 \gev energy range and
W44 was simulated over the two-year time range included in the
2FGL catalog. 
In the simulations, we disabled the inherent energy dispersion of photons
by the LAT. The bias caused by ignoring energy dispersion 
will be discussed in an upcoming paper by the LAT collaboration \citep{lat_on_orbit_psf}.
In total, we performed 985 identical simulations.

@@

Figure~\ref{ts_comparison_w44sim}a shows the distribution of \ts values when 
fitting the simulations of W44 as as a point-like source.
The sources are very significant ($\ts\sim3500$).
Figure~\ref{ts_comparison_w44sim}b shows the distribution of \tsext values
when fitting the source as a radially-symmetric uniform disk and Gaussian source.
The sources are very significantly extended ($\tsext\sim250$). 
On average \tsext is slightly larger when
fitting the source with the disk spatial model compared to the Gaussian spatial model,
consistent with the fact that the simulated source's true spatial model has a sharp edge.
Figure~\ref{ts_comparison_w44sim}c then shows the change in \ts 
when fitting the simulated source as
an elliptical disk compared to fitting the source as a disk. Since the changes
in \ts are on average $\sim30$, we see that our simulated source is significantly better
described with an elliptical spatial model. Finally, 
Figure~\ref{ts_comparison_w44sim}d shows the changes in \ts when fitting
the sources with an elliptical ring spatial model compared to fitting them
with elliptical disk spatial models.
Since the changes are on average $\sim9$, the simulated sources are only somewhat better described
by the true spatial model.

Figure~\ref{bias_w44sim}a and Figure~\ref{bias_w44sim}b
show the distributions of best fit flux and spectral indices when fitting
the simulated sources.
When fitting the sources with a point-like
spatial model, there is a $\sim$10\% bias in the fit flux and
a $\sim5\%$ bias in the fit index. When fitting the sources assuming any
of the extended spatial models, there is no bias in the best fit spectral parameters
(within the precision of the Monte Carlo study). The reason is that
within the finite angular precision and limited statistics of the LAT, 
all of the extended spatial models can be adjusted to look very similar.

Figure~\ref{bias_w44sim}c plots the best fit 68\% containment radius.  
For the elliptical ring spatial model, there is no significant bias in the best fit size. 
None of the other spectral models induced a bias greater
than $\sim10\%$. From this, we see that \rsixeight is a reasonable measure of the true 68\%
containment radius for a reasonable spatial models.

For completeness, Figure~\ref{bias_w44sim}d plots the distribution of best
fit elliptical eccentricities and Figure~\ref{bias_w444im}e plots the
distribution of orientation angles
when the sources were fit with elliptical
disk and elliptical ring spatial models.  
The eccentricity is defined as
$\sqrt{1-b^2/a^2}$, where $a$ is the length of the semi-major axis and
$b$ is the length of the semi-minor axis.  
There is no systematic bias in the eccentricities or orientation angles.
Finally, Figure~\ref{bias_w44sim}f shows that the distribution of best fit
ring fractions is unbiased when the sources are fit with the elliptical
ring spatial model.
}

@@

\newtext{
We considered an alternate Bayesian formulation of this test.
We considered the Bayesian information criterion \citep[BIC,][]{BIC_statistical_test},
but found it difficult
to apply to LAT data because it required a term including the number of data points
and for LAT events. Because there are many photons included in an analysis that contribute only
 marginally to the analysis of a source, it is unclear how this test could be applied.
We also considered the Akaike information criterion test \citep[\aic,][]{AIC_statistical_test}.
The \aic is defined as
\begin{equation}
  \aic=2k-2\log\likelihood,
\end{equation}
where $k$ is the number of parameters in the model. 
In this formulation, the best model is considered to be the one that minimizes the \aic.
The first term penalizes models with additional parameters.

Because the two point-like source hypotheses has
three more parameters than the single extended source hypothesis (two spatial
parameters and two spectral parameters compared to one extension parameter),
we note that the $\tsext>\tsinc$ is mathematically equivalent to the test:
\begin{equation}
  \aic_\text{ext} < \aic_\text{2pts} - 6.
\end{equation}
Therefore, our original test has an 
equivalent Bayesian formulation requiring that $\aic_\text{ext}$ 
be sufficiently smaller than $\aic_\text{2pts}$.
}

@@

2FGL\,J1632.4$-$4753c was found by our search to be an extended source
candidate in the 10 \gev to 100 \gev energy range but is in a crowded
region of the sky.  It is spatially coincident with the \tev source
HESS\,J1632$-$478.  Figure~\ref{1FGL_J1632.9-4802c}a shows a counts
map of this source and overlays \tev contours of HESS\,J1632$-$478
\citep{hess_plane_survey}.  There are six nearby point-like 2FGL
sources that appear to represent physically distinct sources
and were included in our background model: 2FGL\,J1630.2$-$4752, 2FGL\,J1631.7$-$4720c, 2FGL\,J1632.4$-$4820c, 2FGL\,J1635.4$-$4717c,
2FGL\,J1636.3$-$4740c, and
2FGL\,J1638.0$-$4703c. On the
other hand, one point-like 2FGL source 2FGL\,J1634.4$-$4743c  correlates
with the extended \tev source and at \gev energies does
not appear physically separate. 
\newtext{It is very close to the position of 2FGL\,J1632.4$-$4753c
and does not show in the observed photon distribution separated emission.}
We therefore removed this source from our model of the
background.  Figure~\ref{1FGL_J1632.9-4802c}b shows the same region with
the background sources subtracted.  With this model, 2FGL\,J1632.4$-$4753c
was found to have an extension $\sigma=0\fdg35\pm0\fdg04_\stat\pm0\fdg02_\sys$
with $\tsext=26.9$.  While fitting the
extension of 2FGL\,J1632.4$-$4753c, we iteratively relocalized
2FGL\,J1635.4$-$4717c to $(l,b)=(337\fdg23,0\fdg35)$ and
2FGL\,J1636.3$-$4740c to $(l,b)=(336\fdg97,-0\fdg07)$.

@@

\tablenotetext{(a)}{Integral 100 \mev to 100 \gev flux.}
\tablenotetext{(b)}{
\newtext{For the Galactic simulations, the quoted fluxes are
if the sources were placed in the Galactic center and the actual
flux is scaled by Equation~\ref{scale_flux_by_background}.}
}

\tablecomments{
    A list of the spectral models of the simulated point-like sources
    which were tested for extension.  For each model, the number of
    statistically independent simulations and the average value of \ts is
    also tabulated.  
    \newtext{The top rows are the sources simulated on top of an isotropic background and
    the bottom rows are the simulations on top of the Galactic diffuse and isotropic
    background.}
}
\end{deluxetable}

@@

\clearpage
\begin{figure}
    \ifcolorfigure
    \plotone{mc_plots/ts_comparison_w44sim_color.eps}
    \else
    \plotone{mc_plots/ts_comparison_w44sim_bw.eps}
    \fi
    % this plot came from $w44simplots/plot_ts_values.py
    \caption{
    \newtext{The distribution of \ts values when fitting statistically
    independent simulations of W44. (a) is the distribution of \ts values
    when fitting W44 as a point-like source and (b) is the
    distribution of \tsext when fitting the source with a Gaussian and uniform disk 
    spatial model. (c) is the distribution of the change in TS when
    fitting the source with an elliptical disk spatial model compared to
    fitting it with a radially-symmetric disk spatial model and (d) 
    when fitting the source with a elliptical ring spatial model compared
    to an elliptical disk spatial model.
    }
    }\label{ts_comparison_w44sim}
\end{figure}

>>>>

 For these reasons, in Section~\ref{analysis_methods_section}
we present a new systematic method for analyzing spatially extended
LAT sources.  
\newtext{
In Section~\ref{monte_carlo_validation}, we demonstrate
that this method can be used to test the statistical significance of the
extension of a LAT source and in Section~\ref{bias_wrong_spatial_model}
we asses the bias that is introduced by using an incorrect spatial model.
In Section~\ref{extension_sensitivity},
we calculate the LAT detection threshold to resolve the extension
of a source.}
In Section~\ref{dual_localization_method}, we
study the ability of the LAT to 
distinguish between a single extended source and unresolved closely-spaced point-like sources
In Section~\ref{test_2lac_sources}, we further demonstrate that our
detection method does not misidentify point-like sources as being
extended by testing the extension of active Galactic nuclei (AGN)
believed to be unresolvable.  In Section~\ref{validate_known},
we systematically reanalyze the twelve extended sources included
in the 2FGL catalog and in Section~\ref{systematic_errors_on_extension}
we describe a way to estimate systematic errors on the measured extension of a source.
In Section~\ref{extended_source_search_method}, we
describe a search for new spatially extended LAT sources. Finally,
in Section~\ref{new_ext_srcs_section} we present the detection of the
extension of nine spatially extended sources that were reported in the 2FGL catalog
but treated as point-like in the analysis.  Two of these sources have been previously analyzed in
dedicated publications.

@@

For sources that are comparable in size to the PSF,
the differences in the PDF for
different spatial models are lost in the noise and the LAT is not sensitive
to the detailed spatial structure of these sources.  
\newtext{In section \ref{bias_wrong_spatial_model}, we perform a dedicated Monte Carlo simulation
to asses possible bias due to incorrectly modeling the spatial structure
of an extended source.
In our search for extended sources, we use only a radially-symmetric uniform
disk spatial model. Unless otherwise noted,
we quote the radius to the edge ($\sigma$) as the size of the source.}

@@

We evaluated the false detection probability for the measurement of source extent
by fitting the extensions of point-like sources.  \cite{mattox_egret}
discuss that the TS distribution for a likelihood ratio test
on the existence of a source at a given position is
\begin{equation}\label{ts_ext_distribution}
  P(\ts)=\onehalf(\chi^2_1(\ts)+\delta(\ts)),
\end{equation}
\newtext{
where $P(\ts)$ is the probability density to get a particular value of TS,
$\chi^2_1$ is the chi-squared distribution with one degree of freedom, and
$\delta$ is the Dirac delta function.
}
The particular form of Equation \ref{ts_ext_distribution} is due to the
null hypothesis (source flux $\Phi=0$) residing on the edge of parameter
space and the model hypothesis adding a single degree of freedom.
It leads to the often quoted result $\sqrt{TS}=\sigma$, where 
$\sigma$ here refers to the significance of the detection. It is plausible
to expect a similar distribution for the TS in the test for
source extension since the same conditions apply (with the source flux
$\Phi$ replaced by the source radius $r$ and $r<0$ being unphysical).
To verify Equation~\ref{ts_ext_distribution}, we evaluated the
empirical distribution function of \tsext computed from simulated sources.

@@

\newtext{
We performed a second simulation to assess any effects that could
be introduces by performing this test in the presence of a highly
structured background.
Our simulation setup was the same as above except that the sources were
simulated on top of and analyzed assuming the presence of the standard
Galactic diffuse and isotropic background models used in 2FGL.  In our
simulations, we selected our sources to have a random position on the sky
such that the absolute value of the galactic latitude of the source was
less than 5\degree. This probes the most structured areas of the 
Galactic background.

To limit the number of tests, we selected only one flux
level for each of the four spectral indices and we perform
this test only in the 1 \gev to 100 \gev energy range. 
The fluxes were selected so that $\ts\sim50$. We do not
expect to be able to spatially resolve sources much dimmer than this
and sources much brigher are not likely to be affected by the background.

Because the Galactic diffuse emission is highly structure, the point-source
sensitivity can vary significantly across the galactic plane.
To account for this, we scaled the flux (for a given spectral index)
so that the source always has approximately the same signal-to-noise ratio:
\begin{equation}
  \label{scale_flux_by_background}
  F(\vec{x}) = F(\text{GC}) \times \left(
  \frac{B(\vec{x})}{B(\text{GC})}\right)^{1/2}.
\end{equation}
Here, $\vec{x}$ is the position of the simulated source, $F$ is the integral
flux of the source from 100 \mev to 100 \gev, $F(\text{GC})$
is the same quantity if the source were at the galactic center, $B$
is the integral of the Galactic diffuse and isotropic emission
from 1 \gev to 100 \gev, and $B(\text{GC})$ is the same quantity
if the source was at the galactic center.  Table~\ref{ts_ext_num_sims}
contains $F(\text{GC})$ and the average value of \ts for the four spectral models.

For each spectrum, we performed $\sim90,000$ simulations.
Figure~\ref{tsext_plane_plot} shows the cumulative density
of \tsext. For small values of \tsext,
there is good agreement between the simulations and
theory.  For the highest values of \tsext, there is possibly
a small discrepancy, but limited statistics makes this hard to quantify.
}

@@

\newtext{
We performed a Monte Carlo study to quantify the bias that is introduced by
incorrectly modeling the spatial structure of an extended source. 
To assess this systematic, we simulated the spatially extended ring-type SNR W44.
We selected W44 because it is the most significant extended source detected by the LAT 
that has a non-radially symmetric photon-distribution \citep{w44}.

W44 was simulated with a power-law spectral model with an integral flux
of $7.12\times10^{-8}$ \phflux in the energy range from 1 \gev to 100
\gev and a spectral index of 2.66.
The simulated flux and spectral index of this source
were taken from our analysis of W44 which is described in Section~\ref{validate_known}.

W44 was simulated with the elliptical ring spatial model described in
\cite{w44}. For reference, the ellipse has a semi-major axis of 0\fdg3,
a semi-minor axis of 0\fdg19, has a position angle of $147\degree$
measured east of celestial north, and predicts emission only for the
outer 0.75\% of the shape.

We used a simulation setup similar to that described in
Section~\ref{monte_carlo_validation}, but the source was simulated on
top of and analyzed assuming the presence of the standard Galactic diffuse and
isotropic background model used in 2FGL.  The simulations were
in the 1 \gev to 100 \gev energy range and
W44 was simulated over the two-year time range used in 2FGL. 
In the simulations, we disabled the inherent energy dispersion of photons
by the LAT. The bias caused by ignoring energy dispersion 
will be discussed in an upcoming paper by the LAT collaboration \citep{lat_on_orbit_psf}.
In total, we performed 985 identical simulations.

@@

Figure~\ref{ts_comparison_w44sim}a shows that the simulations of
W44 when fit as point-like sources are very significant ($\ts\sim3500$).
Figure~\ref{ts_comparison_w44sim}b shows that the simulations when
fit with a radially-symmetric uniform disk or Gaussian spatial
model are highly extended ($\tsext\sim250$).  On average \tsext is
slightly larger when fitting the sources with a uniform disk spatial
model compared to a Gaussian spatial model, consistent with the
fact that the simulated source's true spatial model has a sharp edge.
Figure~\ref{ts_comparison_w44sim}c shows that the sources fit better
when assuming an elliptical disk spatial model compared to a uniform disk
spatial model ($\ts_\text{elliptical\ disk}-\ts_\text{disk}\sim30$).
Finally, Figure~\ref{ts_comparison_w44sim}d shows that the sources
are only fit slightly better assuming an elliptical ring spatial model
compared to an elliptical disk spatial model ($\ts_\text{elliptical\
ring}-\ts_\text{elliptical\ disk}\sim9$).

Figure~\ref{bias_w44sim}a and Figure~\ref{bias_w44sim}b
show the distributions of best fit fluxes and spectral indices when fitting
the simulated sources.
When fitting the sources with a point-like
spatial model, there is a $\sim$10\% bias in the fit flux and
a $\sim5\%$ bias in the fit index. When fitting the sources assuming any
of the extended spatial models, there is no bias in the best fit spectral parameters
(within the precision of the Monte Carlo study). This is because,
within the finite angular resolution and limited statistics of the LAT, 
all of the extended spatial models can be adjusted to look very similar.

Figure~\ref{bias_w44sim}c plots the best fit 68\% containment radius when
the sources are fit with the extended spatial models.  When assuming
the elliptical ring spatial model, there is no significant bias in the
best fit size.  Becasue none of the other spectral models induced a
bias greater than $\sim10\%$, we see that \rsixeight for any of the
spatial models can be used as a reasonable measure of the true 68\%
containment radius for the source.

For completeness, Figure~\ref{bias_w44sim}d and Figure~\ref{bias_w44sim}e 
shows that there are 
no systematic biases in the
distribution of best
fit elliptical eccentricities and orientation angles
when the sources were fit with elliptical
spatial models.  
The eccentricity is defined as
$\sqrt{1-b^2/a^2}$, where $a$ is the length of the semi-major axis and
$b$ is the length of the semi-minor axis.  
Finally, Figure~\ref{bias_w44sim}f shows that the distribution of best fit
inner ring fractions is unbiased when the sources are fit with the elliptical
ring spatial model.
}

@@

\newtext{
We considered using 
the Bayesian information criterion \citep[BIC,][]{BIC_statistical_test} as
an alternate Bayesian formulation of this test. But it is difficult to apply
to LAT data because it contains a term including the number of data points. But for LAT data,
there are many photons included in an analysis that contribute only
 marginally to the analysis of a source
As an alternative, we considered the Akaike information criterion test \citep[\aic,][]{AIC_statistical_test}.
The \aic is defined as
\begin{equation}
  \aic=2k-2\log\likelihood,
\end{equation}
where $k$ is the number of parameters in the model. 
In this formulation, the best model is considered to be the one that minimizes the \aic.
The first term penalizes models with additional parameters.

Because the two point-like source hypotheses has
three more parameters than the single extended source hypothesis (two spatial
parameters and two spectral parameters compared to one extension parameter),
we note that the $\tsext>\tsinc$ test is mathematically equivalent to testing 
\begin{equation}
  \aic_\text{ext} < \aic_\text{2pts} - 6.
\end{equation}
Therefore, our test has an 
equivalent Bayesian formulation requiring that $\aic_\text{ext}$ 
be sufficiently smaller than $\aic_\text{2pts}$.
}

@@

2FGL\,J1632.4$-$4753c was found by our search to be an extended source
candidate in the 10 \gev to 100 \gev energy range but is in a crowded
region of the sky.  It is spatially coincident with the \tev source
HESS\,J1632$-$478.  Figure~\ref{1FGL_J1632.9-4802c}a shows a counts
map of this source and overlays \tev contours of HESS\,J1632$-$478
\citep{hess_plane_survey}.  There are six nearby point-like 2FGL
sources that appear to represent physically distinct sources
and were included in our background model: 2FGL\,J1630.2$-$4752, 2FGL\,J1631.7$-$4720c, 2FGL\,J1632.4$-$4820c, 2FGL\,J1635.4$-$4717c,
2FGL\,J1636.3$-$4740c, and
2FGL\,J1638.0$-$4703c. On the
other hand, one point-like 2FGL source 2FGL\,J1634.4$-$4743c  correlates
with the extended \tev source and at \gev energies does
not appear physically separate. 
\newtext{It is very close to the position of 2FGL\,J1632.4$-$4753c
and does not show separated emission in the observed photon distribution.}
We therefore removed this source from our model of the
background.  Figure~\ref{1FGL_J1632.9-4802c}b shows the same region with
the background sources subtracted.  With this model, 2FGL\,J1632.4$-$4753c
was found to have an extension $\sigma=0\fdg35\pm0\fdg04_\stat\pm0\fdg02_\sys$
with $\tsext=26.9$.  While fitting the
extension of 2FGL\,J1632.4$-$4753c, we iteratively relocalized
2FGL\,J1635.4$-$4717c to $(l,b)=(337\fdg23,0\fdg35)$ and
2FGL\,J1636.3$-$4740c to $(l,b)=(336\fdg97,-0\fdg07)$.

@@

\tablenotetext{(a)}{Integral 100 \mev to 100 \gev flux.}
\tablenotetext{(b)}{
\newtext{For the Galactic simulations, the quoted fluxes are
the flux for sources placed 
in the Galactic center. The actual flux is scaled by Equation~\ref{scale_flux_by_background}.}
}

\tablecomments{
    A list of the spectral models of the simulated point-like sources
    which were tested for extension.  For each model, the number of
    statistically independent simulations and the average value of \ts is
    also tabulated.  
    \newtext{The top rows are the simulations on top of an isotropic background and
    the bottom rows are the simulations on top of the Galactic diffuse and isotropic
    background.}
}
\end{deluxetable}

@@

\clearpage
\begin{figure}
    \ifcolorfigure
    \plotone{mc_plots/ts_comparison_w44sim_color.eps}
    \else
    \plotone{mc_plots/ts_comparison_w44sim_bw.eps}
    \fi
    % this plot came from $w44simplots/plot_ts_values.py
    \caption{
    \newtext{The distribution of \ts values when fitting statistically
    independent simulations of W44. (a) is the distribution of \ts values
    when fitting W44 as a point-like source and (b) is the
    distribution of \tsext when fitting the source with a uniform disk and Gaussian
    spatial model. (c) is the distribution of the change in TS when
    fitting the source with an elliptical disk spatial model compared to
    fitting it with a radially-symmetric disk spatial model and (d) 
    when fitting the source with an elliptical ring spatial model compared
    to an elliptical disk spatial model.
    }
    }\label{ts_comparison_w44sim}
\end{figure}

\clearpage
\begin{figure}
    \ifcolorfigure
    \plotone{mc_plots/bias_w44sim_color.eps}
    \else
    \plotone{mc_plots/bias_w44sim_bw.eps}
    \fi
    % this plot came from $w44simplots/plot_bias
    \caption{
    \newtext{the distribution of fit parameters
    for the Monte Carlo simulations of W44.
    The plots show the distribution of 
    best fit (a) flux (b) spectral index (c) 68\% containment
    radius (d) eccentricity (e) elliptical orientation angle
    and (f) inner ring fraction.
    }
    }\label{bias_w44sim}
\end{figure}
